# Charging Naked Tn5 with MEDS (Picelli) {-}

## Notes {-}

This procotol is adapted from Picelli, et al. 2014 [@Picelli2014].

Yield can be stored at -20C.

## Workflow {-}

```{r echo=F}
library(DiagrammeR)

grViz("
digraph Transposome {
  node [shape = oval]
  'Annealed MEDS'; 'Naked Transposase'; '2X Dialysis Buffer'
  node [shape = record]
  'Mix Assembly Reaction'
  'Incubate @ RT'; 'Store @ -20C'; 'Remove Unbound MEDS'
  
  {'Annealed MEDS' '2X Dialysis Buffer' 'Naked Transposase'} -> 'Mix Assembly Reaction'
  'Mix Assembly Reaction' -> 'Incubate @ RT'
  'Incubate @ RT' -> {'Remove Unbound MEDS' 'Store @ -20C'}
  'Remove Unbound MEDS' -> 'Store @ -20C'

}
")

```

## Reagents {-}

* HEPES
* KOH
* NaCl
* EDTA
* DTT
* Triton X-100
* Glycerol
* Pre-annealed MEDS
* Naked Transposase
* DNA-binding Columns or AMPure XP beads (Agencourt)

## Protocol {-}

1. Mix Pre-annealed MEDS Oligos with Tn5.

```{r echo=F}
loading_reaction <- data.frame(Reagent = c("100uM equimolar oligos A/B in TE",
                                           '100% Glycerol',
                                           '2x Tn5 dialysis buffer',
                                           'Tn5, A280=3.0 (1.85 mg/ml)'),
                               Quantity = c(0.125,0.4,0.12,0.36),
                               Unit = rep('Volume',4))

knitr::kable(loading_reaction, caption = 'Tn5 Charging Reaction',
             format = 'pandoc')
```

2. Incubate 60 minutes at room temperature.

3. Charged Tn5 solution can be stored at -20. Unbound MEDS oligos should
be removed from the solution before use. This can be done with a DNA binding column 
([Zymo](https://www.zymoresearch.com/dna-clean-concentrator) recommended in original protocol) or Agencourt AMPure XP beads ([Beckman Coulter](https://www.beckmancoulter.com/wsrportal/wsrportal.portal?_nfpb=true&_windowLabel=UCM_RENDERER&_urlType=render&wlpUCM_RENDERER_path=%252Fwsr%252Fresearch-and-discovery%252Fproducts-and-services%252Fnucleic-acid-sample-preparation%252Fagencourt-ampure-xp-pcr-purification%252Findex.htm#2/10//0/25/1/0/asc/2/A63880///0/1//0/%2Fwsrportal%2Fwsr%2Fresearch-and-discovery%2Fproducts-and-services%2Fnucleic-acid-sample-preparation%2Fagencourt-ampure-xp-pcr-purification%2Findex.htm/)).

## Buffers {-}

### 2x Tn5 Dialysis Buffer {-}

```{r, results='hold', echo=F}
Reagent <- c('HEPES-KOH (pH 7.2)', 'NaCl', 'EDTA', 'DTT', 'Triton X-100',
             'Glycerol', 'H20')

Quantity <- c(1,400,4, 20,20,2,6.556)

# https://stats.idre.ucla.edu/r/codefragments/greek_letters/

Unit <- c("\\(mL\\)","\\(uL\\)","\\(uL\\)","\\(uL\\)","\\(uL\\)","\\(mL\\)","\\(mL\\)")

Final <- c(paste0(100,"\\(m\\)","M"),
           paste0(200,"\\(m\\)","M"),
           paste0(0.2,"\\(m\\)","M"),
           paste0(2,"\\(m\\)","M"),
           '0.2%', '20%','-')

Stock <- c('1M','5M','0.5M','1M','100%','100%','-')

Tn5Dialysis2X <- data.frame(Reagent=Reagent,
                            Quantity=Quantity,
                            Unit=Unit,
                            "Final Conc."=Final,
                            Stock=Stock)
knitr::kable(Tn5Dialysis2X, caption = '2X Tn5 Dialysis Buffer',
             format = 'pandoc')
```
