---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Cell Line ChIP-seq {-}

## Notes {-}

This protocol is adapted from the ChIP protocol used in the Bradner Lab and Christoph Bock's tagmentation v1.1. Works well with 10-25 million cells.
Unless otherwise noted keep all steps on ice. Work quickly and reduce handling of material as much as possible.
For example, never pipette mix when resuspension in a rotator at 4c would suffice.

When using a bioruptor, 0.5% SDS is usually optimal. Qsonica requires 1% SDS.
Before starting, make buffered stocks of HEPES and Tris.

## Workflow {-}
```{r out.height='3.5in', out.width='10in', fig.height=1,fig.width=7, fig.align="center", echo=F}
library(DiagrammeR)
grViz("digraph ChIP {
  'Fixation' -> 'Lysis'
  'Lysis' -> 'Sonication'
  'Sonication' -> 'IP'
  'IP' -> 'Tagment IP' 
  'IP' -> 'Store input'
  'Store input' -> Purification
  IP -> Wash -> Purification
  'Tagment IP' -> Purification
  Purification -> 'Tagment input'
  Purification -> Amplify
  'Tagment input' -> Amplify
}")
```

## Protocol {-}


1. Fixation (Can be performed well in advance)
    + Make 11% formaldehyde fixation solution fresh (see below).
    + Add volume of fixation solution equal to 1/10 of the culture volume.
    + Incubate at RT on rocker for 10-15 minutes (optimize for cell type and IP target).
    + Quench with volume of 2.5M glycine equal to 1/20 culture volume.
    + If adherent:
        + Rinse 2x with chilled PBS+PIC at 1X concentration.
        + Add small volume of PBS+PIC (dependent on dish size) and scrape with cell scrapers or alternatively bent 200ul pipette tip.
        + Transfer scraped cells to conical (50ml or 15ml).
    + Wash:
        + Spin at 4c @ 1350g for 5min.
        + Aspirate supernatant.
        + Resuspend in cold 1X PBS+PIC.
        + Repeat 2x.
    + Collect cells by spinning at 4c @ 1350g for 5min, aspirate supernatant.
    + Cells are now ready for lysis. Alternatively snap freeze by dunking conical into LN~2~ and immediately transfer to -80c.
2. Bead setup (Day 0 or Day 1)
    + Wash 25ul Protein G DynaBeads 3x5min in 1ml blocking buffer using magnetic rack and aspirating all buffer with each wash.
    + Add 500ul blocking buffer to tub with trapped beads
    + Add antibody (usually 5-10ug) to this buffer. Do not bother resuspending manually. Instead rotate end over end at 4c for minimum 5 hrs.
3. Lysis (Day 1)
    + Resuspend pellet in chilled LB1+PIC  (>=1ml/1e6 cells) and incubate 10 minutes @ 4c while rotating.
    + Pellet cells at 1350g for 5min @ 4c, aspirate supernatant.
    + Repeat the previous two steps using LB2+PIC.
    + Resuspend in appropriate volume SB+PIC with SDS at optimized percentage.
4. Sonication (Day 1)
    + Sonicate using parameters optimized for instrument, cell type, and cell number.
    + Spin sonicated material at 20,000g for 10min at 4c to clarify.
    + Collect supernatant and transfer to a new tube. Discard pellet (debris).
    + Set aside ~10% total sonicated volume for use as input (non-IP control) and store at -20c.
    + Dilute remaining sonicated chromatin with SB+PIC such that %SDS < 0.1%.
        * Chromatin can be stored at 4c for several days.
5. IP (Day 1)
    + Wash antibody-bead complex 3x5min in cold blocking buffer.
    + Resuspend in 500ul cold blocking buffer.
    + Gently add bead suspension to sonicated chromatin.
    + Rotate end over end at 4c overnight.
6. Bead Wash (Day 2)
    + Collect bead-antibody-chromatin on magnetic rack, aspirate SB + debris.
    + Wash end over end 5min at 4c in each of the following buffers. After each wash, collect beads with magnet, aspirate.
        * Twice in SB
        * High salt SB
        * LiCl wash buffer
        * TE+NaCl (add 10ul 5M NaCl for every 1ml TE)
    + Remove as much of final wash as possible while beads are trapped by the magnetic rack. 
    + If necessary, spin at 4c at 960g for 3 minutes then return to magnetic rack and aspirate again.
    + At this point, dna can be eluted in EB (see below), crosslinks reversed, RNAse and Proteinase K treated, and then 
    purified via column or phenol chloroform extraction. Alteratively, continue on
    to produce NGS libraries via tagmentation.
7. IP Tagmentation (Day 2)
    + Wash Beads once in 10mM Tris HCl, pH 8.
    + Resuspend beads in 25ul Tagmentation Reaction.
    + Incubate at 37c for up to 10 minutes (optimize if necessary).
    + Add 1ml cold Tagmentation Wash Buffer and incubate 5 minutes on ice.
    + Wash beads
        * 2x in Tagmentation Wash Buffer
        * 2x in TE  
        
```{r echo=F}
# per sample
Stock.Reagent <- c("1X Tagmentation Buffer",
                   "Tn5 Enzyme",
                   "Target-Ab-Bead Conjugate")

vol<- c("24ul","1ul","-")


TagmentationMixture <- data.frame(Ingredient=Stock.Reagent,
                             Vol=vol)

knitr::kable(TagmentationMixture, caption = 'Per sample IP Tagmentation Reaction',
             format = 'pandoc')
```
        
8. Elution (Day 2)
    + Trap beads on magnet and remove TE.
    + Resuspend in 200ul EB and incubate at 65c for 15 minutes, vortexing every 5 minutes.
    + Collect beads by centrifuging @ 20,000g or trap on magnetic rack and transfer to fresh tube.
8. Reverse crosslinks (Day 2)
    + Resuspend in 200ul Elution Buffer.
    + At this point, retrieve input samples and bring to the same volume as IP samples using Elution Buffer.
    + Reverse crosslinks at 65c overnight.
9. Purification (Day 3)
    + Add 200ul TE to each tube.
    + Add RNase A to a final concentration of 0.2mg/ml (8ul if stock is 10mg/ml).
    + Mix gently and incubate at 37c for 2hrs.
    + Add 7ul 300mM CaCl~2~ in 10mM Tris-HCL, pH8.
    + Add Proteinase K to a concentration of 0.2mg/ml (4ul if stock is 20mg/ml).
    + Mix gently and incubate at 55c for 30min.
    + Purify Using Minelute Kit or similar column-based extraction.
10. Input Tagmentation (Day 3)
    + Purified input DNA can be quantified via Qubit.
    + 2.5ng should be set aside for tagmentation.
    + Create master mix of 1:10 diluted Transposase.
    + Create a 5ul reaction for each 2.5ng of input DNA (see per sample recipe below).
    + Incubate 5min at 55c.
    + During incubation prepare 0.6% SDS master mix (60ul 10% SDS in 940ul Nuclease Free H~2~0).
    + Add 1ul SDS mix to each input reaction and incubate 5min @ RT.
    + Bring up to 10ul with Nuclease Free H~2~0.
    + Use Minelute or other column based purification method to purify DNA. Elute in ~10ul.
    + Continue with amplification alongside IP samples.
    
```{r echo=F}
# per sample
Stock.Reagent <- c("2.5ng Eluted input DNA",
                   "0.1X Tn5 Enzyme",
                   "1X Tagmentation Buffer")

vol<- c("-","1ul","Up to 5ul")

TagmentationMixture <- data.frame(Ingredient=Stock.Reagent,
                             Vol=vol)

knitr::kable(TagmentationMixture, caption = 'Per sample Input Tagmentation Reaction',
             format = 'pandoc')
```

11. Amplification (Day 3 or 4)
    + Prepare for each sample a Library Amplification Reaction as specified in the table below.
    + Preamplify in a thermocycler according to the following program:
        1. 72c for 5min
        1. 98c for 30 seconds
        1. 5 cycles:
            1. 98c for 10 seconds
            1. 63c for 30 seconds
            1. 72c for 1 minute
    + Store this preamplified material at 4c.
    + Prepare qPCR side reaction as described in table below.
    + Using a qPCR instrument, perform the following:
        1. 98c for 30 seconds
        1. 20 cycles:
            1. 98c for 10 seconds
            1. 63c for 30 seconds
            1. 72c for 1 min
    + Plot linear Rn vs cycle count and find the cycle count that corresponds to 1/4 the maximum Rn value. 
    his is the number of additional cycles to perform in a standard thermocycler for each sample.
    + In a standard thermocycler, perform the following:
        1. 98c for 30 seconds
        1. *X* cycles:
            1. 98c for 10 seconds
            1. 63c for 30 seconds
            1. 72c for 1 min
    
```{r echo=F}
# per sample
Stock.Reagent <- c("Tagmented DNA",
                   "Nuclease Free H~2~0",
                   "Nextera AD1 25uM",
                   "Nextera AD2.X 25uM",
                   "NEBnext HiFidelity 2X PCR Master Mix")

vol<- c("10ul","10ul","2.5ul", "2.5ul","25ul")

AmpReaction <- data.frame(Ingredient=Stock.Reagent,
                             Vol=vol)

knitr::kable(AmpReaction, caption = '25ul Library Amplification Reaction',
             format = 'pandoc')
```

```{r echo=F}
Stock.Reagent <- c("Tagmented DNA",
                   "Nuclease Free H~2~0",
                   "Nextera AD1 25uM",
                   "Nextera AD2.X 25uM",
                   "100x Sybr Green",
                   "NEBnext HiFidelity 2X PCR Master Mix")

vol<- c("5ul","4.41ul","0.25ul", "0.25ul", "0.09ul","5ul")

SideReaction <- data.frame(Ingredient=Stock.Reagent,
                             Vol=vol)


knitr::kable(SideReaction, caption = 'qPCR Side Reaction',
             format = 'pandoc')
```

12. Library Cleanup (Day 3 or 4)
    + Add 1.8x volume Ampure XP Beads and mix 10x with pipette, then incubate 10min at RT.
    + Trap with magnet for ~10min to ensure full bead engagement with magnet. Remove supernatant.
    + Wash briefly 2x with fresh 80% EtOH and remove EtOH.
    + Dry with open caps for approx. 5min. Can be placed in incubator for additional speed.
    + Remove from magnets and resuspend beads in 50ul Nuclease Free H~2~0.
    + Incubate 10 minutes @ RT.
    + Place back on magnet and trap beads for 5min, then carefully place supernatant in fresh 1.7ml tube.
13. Size Selection (Day 3 or 4)
    + Add 32.5ul (65% of 50ul) Ampure XP Beads, mix 10x with pipette and incubate 10min @ RT.
    + Trap beads on magnet for 10min, then transfer supernatant to fresh tube, discarding the beads.
    + Add 12.5ul of Ampure XP Beads, mix 10x by pipette, and incubate 10min @ RT.
    + Trap beads with magnet for 10min and discard supernatant.
    + Wash beads twice with fresh 80% EtOH and remove EtOH.
    + Dry with caps open 5min on magnet.
    + Remove from magnet and resuspend beads in 15ul Nuclease Free H~2~0. Incubate 10min.
    + Return tubes to magnet rack and trap beads for 5min.
    + Transfer supernatant to new tube. This is the final library.

## Buffers {-}

Buffers are shown as recipes for a reasonable volume and use reagents
listed at the end of this document in the Stock Solutions section.

```{r echo=F}
# 50ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8", "37% Formaldehyde","ddH~2~O")

target_conc <- c("50mM","100mM","1mM","0.5mM","11%",NA)

amt <- c("2.5ml","1ml", "100ul", "50ul", "14.9ml","31.5ml")

FixationBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(FixationBuffer, caption = '50mL 11% Formaldehyde Fixation Buffer',
             format = 'pandoc')
```

```{r echo=F}
# 50ml
Stock.Reagent <- c("BSA","1X PBS")

target_conc <- c("0.5%",NA)

amt <- c("125mg","25mL")

FixationBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(FixationBuffer, caption = '25mL Blocking Buffer',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "glycerol","IGEPAL CA-630","Triton X-100","PIC","ddH~2~O")

target_conc <- c("50mM","140mM","1mM","10%","0.5%","0.25%","See product",NA)

amt <- c("12.5mL","7mL","0.5mL","25mL","1.25mL","625uL","See product","203mL")

LysisBuffer1 <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(LysisBuffer1, caption = '250 mL Lysis Buffer 1 (LB1)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M TRIS-HCL, pH8","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8","PIC","ddH~2~O")

target_conc <- c("10mM","200mM","1mM","0.5mM","see product",NA)

amt <- c("2.5mL","10mL","0.5mL","0.25mL","see product","236.75mL")

LysisBuffer2 <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(LysisBuffer2, caption = '250 mL Lysis Buffer 2 (LB2)',
             format = 'pandoc')
```


```{r echo=F}
# 250ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8","Triton X-100","10% Na-Deoxycholate","10% SDS","PIC","ddH~2~O")

target_conc <- c("50mM","140mM","1mM","1mM","1%","0.1%","optimize","see product",NA)

amt <- c("12.5mL","7mL","0.5mL","0.5mL","2.5mL","2.5mL","optimize","see product","vol up to 250mL")

SonicationBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(SonicationBuffer, caption = '250 mL Sonication Buffer (LB)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8","Triton X-100","10% Na-Deoxycholate","10% SDS","PIC","ddH~2~O")

target_conc <- c("50mM","500mM","1mM","1mM","1%","0.1%","0.1%","see product",NA)

amt <- c("12.5mL","25mL","0.5mL","0.5mL","2.5mL","2.5mL","2.5mL","see product","vol up to 250mL")

SonicationBufferHS <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(SonicationBufferHS, caption = '250 mL Sonication Buffer High Salt (SBHS)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M Tris-HCL, pH8",
                   "0.5M EDTA, pH 8",
                   "5M LiCl",
                   "IGEPAL CA-630",
                   "10% Na-deoxycholate",
                   "ddH~2~O")

target_conc <- c("20mM","1mM","250mM","0.5%","0.5%",NA)

amt <- c("5mL","0.5mL","12.5mL","1.25mL","12.5mL","218.25mL")

LiClBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(LiClBuffer, caption = '250 mL LiCl Wash Buffer (LWB)',
             format = 'pandoc')
```



```{r echo=F}
# 250ml
Stock.Reagent <- c("1M Tris-HCL, pH8",
                   "0.5M EDTA, pH 8",
                   "10% SDS",
                   "ddH~2~O")

target_conc <- c("50mM","10mM","1%",NA)

amt <- c("12.5mL","5mL","25mL","207.5mL")

ElutionBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(ElutionBuffer, caption = '250 mL Elution Buffer (EB)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M Tris-HCL, pH8",
                   "5M NaCl",
                   "0.5M EDTA, pH 8",
                   "10% SDS",
                   "Na-Deoxycholate",
                   "Triton X-100",
                   "ddH~2~O")

target_conc <- c("10mM","140mM","1mM","0.1%","0.1%","1%",NA)

amt <- c("2.5mL","7.5mL","0.5mL","2.5mL","2.5mL","2.5mL","232mL")

ElutionBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(ElutionBuffer, caption = '250 mL Tagmentation Wash Buffer',
             format = 'pandoc')
```

## Other Reagents + equipment {-}

* Bovine Serum Albumin
* Triton X-100
* 1X PBS
* 37% formaldehyde (with methanol)
* glycerol
* IGEPAL CA630
* Sonicator and associated attachments
* TPX tubes (if sonicating in BioRuptor)
* Cell scrapers (alternatively bend a sterile pipette tip to a 90~o~ angle)
* DynaBeads (usually protein G)
* Desired antibody
* TE Buffer
* Protease inhibitors (Roche tablets or HALT cocktail; add phosphatase inhibitors for phosphorylated marks such as pSer polII)
* Proteinase K
* RNAse A
* Magnetic Racks for 1.5ml eppendorf and optionally 15ml conical
* Heat block, incubator, or water bath
* Nuclease-free H~2~O
