# Cell Line ChIP (Bradner Lab protocol) {-}

## Notes {-}

This protocol is adapted from the ChIP protocol used in the Bradner Lab. Works well with 10-25 million cells.
Unless otherwise noted keep all steps on ice. Work quickly and reduce handling of material as much as possible.
For example, never pipette mix when resuspension in a rotator at 4c would suffice.

When using a bioruptor, 0.5% SDS is usually optimal. 
Before starting, make buffered stocks of HEPES and Tris.

## Workflow {-}
```{r out.height='3.5in', out.width='10in', fig.height=1,fig.width=7, fig.align="center", echo=F}
library(DiagrammeR)
grViz("
digraph ChIP {
  node [shape = oval]
  'Cultured Cells';  
  node [shape = record]
  'Fixation'; 'Lysis'; 'Sonication'; 'IP';'Wash + Purification'
  
  'Cultured Cells' -> 'Fixation'
  'Fixation' -> 'Lysis'
  'Lysis' -> 'Sonication'
  'Sonication' -> 'IP'
  'IP' -> 'Wash + Purification'
}
")
```

## Protocol {-}

1. Bead setup
    + Wash 100ul Protein G DynaBeads 3x5min in 1ml blocking buffer using magnetic rack and aspirating all buffer with each wash.
    + Add 500ul blocking buffer to tub with trapped beads
    + Add antibody (usually 10ug) to this buffer. Do not bother resuspending manually. Instead rotate end over end at 4c for minimum 5 hrs.
2. Fixation
    + Make 11% formaldehyde fixation solution fresh (see below).
    + Add volume of fixation solution equal to 1/10 of the culture volume.
    + Incubate at RT on rocker for 10-15 minutes (optimize for cell type and IP target).
    + Quench with volume of 2.5M glycine equal to 1/20 culture volume.
    + If adherent:
        + Rinse 2x with chilled PBS+PIC at 1X concentration.
        + Add small volume of PBS+PIC (dependent on dish size) and scrape with cell scrapers or alternatively bent 200ul pipette tip.
        + Transfer scraped cells to conical (50ml or 15ml).
    + Wash:
        + Spin at 4c @ 1350g for 5min.
        + Aspirate supernatant.
        + Resuspend in cold 1X PBS+PIC.
        + Repeat 2x.
    + Collect cells by spinning at 4c @ 1350g for 5min, aspirate supernatant.
    + Cells are now ready for lysis. Alternatively snap freeze by dunking conical into LN~2~ and immediately transfer to -80c.
3. Lysis
    + Resuspend pellet in 5ml chilled LB1+PIC and incubate 10 minutes @ 4c while rotating.
    + Pellet cells at 1350g for 5min @ 4c, aspirate supernatant.
    + Repeat the previous two steps using LB2+PIC.
    + Resuspend in appropriate volume SB+PIC with SDS at optimized percentage.
4. Sonication
    + Sonicate using parameters optimized for instrument and cell type.
    + Spin sonicated material at 20,000g for 10min at 4c to clarify.
    + Collect supernatant and transfer to a new tube. Discard pellet (debris).
    + Set aside ~50ul for use as input (non-IP control) and store at -20c.
    + Dilute remaining sonicated chromatin with SB+PIC such that %SDS < 0.1%, ideally targeting < 15ml total volume.
5. IP
    + Wash antibody-bead complex 3x5min in cold blocking buffer.
    + Resuspend in 500ul cold blocking buffer.
    + Gently add bead suspension to sonicated chromatin.
    + Rotate end over end at 4c overnight.
6. Bead Wash + Elution
    + Collect bead-antibody-chromatin on magnetic rack, aspirate SB + debris.
    + Wash end over end 5min at 4c in each of the following buffers. After each wash, collect beads with magnet, aspirate.
        * Twice in SB
        * High salt SB
        * LiCl wash buffer
        * TE+NaCl (add 10ul 5M NaCl for every 1ml TE)
    + Remove as much of final wash as possible while beads are trapped by the magnetic rack. 
    If necessary, spin at 4c at 960g for 3 minutes then return to magnetic rack and aspirate again.
    + Gently resuspend in 210ul EB. Pipette mix vigorously or vortex.
    + Elute chromatin by incubating at 65c for 15min. Vortex or pipette mix once every 5min.
    + Spin at 20,000g to pellet beads and/or trap with magnetic rack.
    + Transfer supernatant to fresh tube
    + If input chromatin was saved, add EB to input chromatin such that IP and input are in the same volume.
    + Reverse crosslinks overnight, ~16hrs at 65c.
6. Purification
    + Add 200ul TE to each tube.
    + Add RNAseA to a concentration of 0.2mg/ml to each tube (8ul if stock is 10mg/ml).
    + Mix gently then incubate at 37c for 2hrs.
    + Add 7ul 300mM CaCl in 10mM ph8 Tris-HCL.
    + Add proteinase K to a concentration of 0.2mg/ml (4ul if stock concentration is 20mg/ml).
    + Mix and incubate 30min at 55c.
    + Options:
        1. Proceed with phenol:chloroform:isoamyl extraction
        2. Use column based extraction (e.g. minelute).

## Buffers {-}

Buffers are shown as recipes for a reasonable volume and use stock solutions and/or reagents listed below.

```{r echo=F}
# 50ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8", "37% Formaldehyde","ddH~2~O")

target_conc <- c("50mM","100mM","1mM","0.5mM","11%",NA)

amt <- c("2.5ml","1ml", "100ul", "50ul", "14.9ml","31.5ml")

FixationBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(FixationBuffer, caption = '50mL 11% Formaldehyde Fixation Buffer',
             format = 'pandoc')
```

```{r echo=F}
# 50ml
Stock.Reagent <- c("BSA","1X PBS")

target_conc <- c("0.5%",NA)

amt <- c("50mg","50mL")

FixationBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(FixationBuffer, caption = '50mL Blocking Buffer',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "glycerol","IGEPAL CA-630","Triton X-100","PIC","ddH~2~O")

target_conc <- c("50mM","140mM","1mM","10%","0.5%","0.25%","See product",NA)

amt <- c("12.5mL","7mL","0.5mL","25mL","1.25mL","625uL","See product","203mL")

LysisBuffer1 <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(LysisBuffer1, caption = '250 mL Lysis Buffer 1 (LB1)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M TRIS-HCL, pH8","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8","PIC","ddH~2~O")

target_conc <- c("10mM","200mM","1mM","0.5mM","see product",NA)

amt <- c("2.5mL","10mL","0.5mL","0.25mL","see product","236.75mL")

LysisBuffer2 <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(LysisBuffer2, caption = '250 mL Lysis Buffer 2 (LB2)',
             format = 'pandoc')
```


```{r echo=F}
# 250ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8","Triton X-100","10% Na-Deoxycholate","10% SDS","PIC","ddH~2~O")

target_conc <- c("50mM","140mM","1mM","1mM","1%","0.1%","optimize","see product",NA)

amt <- c("12.5mL","7mL","0.5mL","0.5mL","2.5mL","2.5mL","optimize","see product","vol up to 250mL")

SonicationBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(SonicationBuffer, caption = '250 mL Sonication Buffer (LB)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M HEPES-KOH, pH7.5","5M NaCl","0.5M EDTA, pH 8",
                   "0.5M EGTA, pH 8","Triton X-100","10% Na-Deoxycholate","10% SDS","PIC","ddH~2~O")

target_conc <- c("50mM","500mM","1mM","1mM","1%","0.1%","0.1%","see product",NA)

amt <- c("12.5mL","25mL","0.5mL","0.5mL","2.5mL","2.5mL","2.5mL","see product","vol up to 250mL")

SonicationBufferHS <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(SonicationBufferHS, caption = '250 mL Sonication Buffer High Salt (LBHS)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M Tris-HCL, pH8",
                   "0.5M EDTA, pH 8",
                   "5M LiCl",
                   "IGEPAL CA-630",
                   "10% Na-deoxycholate",
                   "ddH~2~O")

target_conc <- c("20mM","1mM","250mM","0.5%","0.5%",NA)

amt <- c("5mL","0.5mL","12.5mL","1.25mL","12.5mL","218.25mL")

LiClBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(LiClBuffer, caption = '250 mL LiCl Wash Buffer (LWB)',
             format = 'pandoc')
```

```{r echo=F}
# 250ml
Stock.Reagent <- c("1M Tris-HCL, pH8",
                   "0.5M EDTA, pH 8",
                   "10% SDS",
                   "ddH~2~O")

target_conc <- c("50mM","10mM","1%",NA)

amt <- c("12.5mL","5mL","25mL","207.5mL")

ElutionBuffer <- data.frame(Ingredient=Stock.Reagent,
                             Target.Conc=target_conc,
                             To.Add=amt)

knitr::kable(ElutionBuffer, caption = '250 mL Elution Buffer (EB)',
             format = 'pandoc')
```

## Stocks {-}

```{r echo=F}
Stock <- c("1M HEPES-KOH, pH 7.5","5M NaCl","1M Tris-HCL, pH 8","5M LiCl","10% Na-Deoxycholate","10% SDS","1M (~10%) CaCl~2~", "2.5M Glycine")
target_conc <- c(paste0(1,"M"),
           paste0(5,"M"),
           paste0(1,"M"),
           paste0(5,"M"),
           paste0(10,"%"),
           paste0(10,"%"),
           paste0(1,"M"), paste0(2.5,"M"))

target_vol <- c(paste0(200,"mL"),
                paste0(500,"mL"),
                paste0(200,"mL"),
                paste0(40,"mL"),
                paste0(200,"mL"),
                paste0(200,"mL"),
                paste0(100,"mL"),
                paste0(400,"mL"))
mw <- c(238.3,58.44,121.14,42.39,NA,NA, 110.98, 75.07)

solid=c(paste0(47.66,"g"),
        paste0(146.1,"g"),
        paste0(24.228,"g"),
        paste0(8.478,"g"),
        paste0(20,"g"),
        paste0(20,"g"),
        paste0(11.1,"g"),
        paste0(75.07,"g"))

adjustment <- c("KOH","-","HCL","-","-","-","-","-")

ChIPStocks <- data.frame(Stock=Stock, 
                         Target.Vol=target_vol, 
                         Target.Conc=target_conc, 
                         Reagent.MW=mw, 
                         solid=solid, Adjustment = adjustment)

knitr::kable(ChIPStocks, caption = 'ChIP stock solutions',
             format = 'pandoc')
```

## Other Reagents + equipment {-}

* Bovine Serum Albumin
* Triton X-100
* 1X PBS
* 37% formaldehyde (with methanol)
* glycerol
* IGEPAL CA630
* Sonicator and associated attachments
* TPX tubes (if sonicating in BioRuptor)
* Cell scrapers (alternatively bend a sterile pipette tip to a 90~o~ angle)
* DynaBeads (usually protein G)
* Desired antibody
* TE Buffer
* Protease inhibitors (Roche tablets or HALT cocktail; add phosphatase inhibitors for phosphorylated marks such as pSer polII)
* Proteinase K
* RNAse A
* Magnetic Racks for 1.5ml eppendorf and optionally 15ml conical
* Heat block, incubator, or water bath
* Nuclease-free H~2~O
